Hashlattice: A distributed system for monotonic peer-to-peer synchronization

### 2022-11-20 paper/hl_paper.tex
semilattice graphs

>>>>

    \subsection{Semilattice graphs}

      We can extend the notion of dependent semilattices to a graph of semilattices depending on each other.

      As a pre-requisite, we define a semilattice over $\mathrm{Option}(L)$ given a semilattice over $L$ as follows:

      A \emph{semilattice graph} consists of:
      \begin{itemize}
        \item a totally ordered set $K$ of keys, representing semilattices that may depend on previous semilattices
        \item a set $V$ of values of any of these semilattices
        \item a function $\checkelem : \prod_{k \in K} (K_{<k} \to \mathrm{Option}(V)) \times V \to \mathrm{Bool}$ assining to each $K$ and corresponding semilattice value a boolean indicating whether the value is an element of the semilattice, which depends on only previous semilattice elements in a way continuous in the $K_{<k} \to \mathrm{Option}(V)$ argument under the product topology.
        \item a function $\join : \prod_{k \in K} (K_{<k} \to \mathrm{Option}(V)) \times V \times V \to V$ assigning to each $K$ and corresponding semilattice values a join, which depends on only previous semilattice elements in a way continuous in the $K_{<k} \to \mathrm{Option}(V)$ argument under the product topology.
        \item a function $\transport : \prod_{k \in K} (K_{<k} \to \mathrm{Option}(V)) \times (K_{<k} \to \mathrm{Option}(V)) \times V \to \mathrm{Option}(V)$ assigning to each $K$ and semilattice value a transported value, which depends on previous semilattice elements in both the previous and new contexts, in a way continuous in the $K_{<k} \to \mathrm{Option}(V)$ arguments under the product topology.
      \end{itemize}

      These must satisfy dependency conditions and dependent semilattice conditions. The dependency conditions state that there exists some function $d : \prod{k \in K} (K_{<k} \to \mathrm{Option}(V)) \times V \rightarrow 2^K$ such that:

      \begin{itemize}
        \item For $k : K, c : (K_{<k} \to \mathrm{Option}(V)), v : V$, the function $c' \mapsto \checkelem(k, c', v)$ equals $\checkelem(k, c, v)$ if $c'(k') = c(k')$ for all $k' \in d(k, c, v)$.
        \item For $k : K, c : (K_{<k} \to \mathrm{Option}(V)), v_1, v_2 : V$, the function $c' \mapsto \join(k, c', v_1, v_2)$ equals $\join(k, c, v_1, v_2)$ if $c'(k') = c(k')$ for all $k' \in d(k, c, v_1) \cup d(k, c, v_2)$.
        \item For $k : K, c_1 : (K_{<k} \to \mathrm{Option}(V)), c_2 : (K_{<k} \to \mathrm{Option}(V)), v : V$, the function $(c_1', c_2') \mapsto \transport(k, c_1', c_2', v)$ equals $\transport(k, c_1, c_2, v)$ if $c_1'(k') = c_1(k')$ and $c_2'(k') = c_2(k')$ for all $k' \in d(k, c_1, v) \cup d(k, c_2, v)$.
      \end{itemize}

      To define the dependent semilattice conditions, we first define a partial order over partial key-value mappings. If $k : K, v : V, c_1, c_2 : (K_{<k} \to \mathrm{Option}(V))$, if $\checkelem(k, c_1, v)$ and $\checkelem(k, c_2, v)$, then:

      $$c_1 \leq_{k, v} c_2 \Leftrightarrow \forall k' \in d(k, c_1) \cup d(k, c_2), c_1(k') \leq c_2(k')$$.

      The dependent semilattice conditions state that:

      \begin{itemize}
        \item For $k : K, c : (K_{<k} \to \mathrm{Option}(V))$, then $(v_1 : V, v_2 : V) \mapsto \join(k, c, v_1, v_2)$ is the join of a semilattice over $\{ v : V ~ | ~ \checkelem(k, c, v) \}$.
        \item For $k : K, c_1, c_2 : (K_{<k} \to \mathrm{Option}(V)), v_1 : V$, if $\checkelem(k, c_1, v_1)$ and $c_1 \leq_{k, v_1} c_2$, then $\transport(k, c_1, c_2, v_1) = \mathrm{Some}(v_2)$, then $\checkelem(k, c_2, v_2)$.
        \item For $k : K, c : (K_{<k} \to \mathrm{Option}(V)), v : V$, if $\checkelem(k, c, v)$, then $\transport(k, c, c, v) = \mathrm{Some}(v)$.
        \item For $k : K, c_1, c_2, c_3 : (K_{<k} \to \mathrm{Option}(V)), v_1 : V$, if $\checkelem(k, c_1, v_1)$ and $c_1 \leq_{k, v_1} c_2$ and $\transport(k, c_1, c_2, v_1) = \mathrm{Some}(v_2)$ and $c_2 \leq_{k, v_2} c_3$, then $\transport(k, c_1, c_3, v_1) = \transport(k, c_2, c_3, v_2)$.
        \item For $k : K, c_1, c_2 : (K_{<k} \to \mathrm{Option}(V)), v_1, v_1' : V$, if $\checkelem(k, c_1, v_1)$ and $\checkelem(k, c_1, v_1')$, then $\transport(k, c_1, c_2, \join(k, c_1, v_1, v_1')) = \transport(k, c_1, v_1) \vee_{k, c_2} \transport(k, c_1, v_1')$, where $\vee_{k, c_2}$ is the join of the option semilattice implied by the join $(v, v') \mapsto \join(k, c_2, v, v')$.
      \end{itemize}
